Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-20 ba_resubmit/response_to_reviewers2.tex

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

%% RCS: Suggestion: Write a strong paragraph to the Editor after the paper is finalized. Come back to this. Explain that we have tried to simplify the notation when possible, however, in some places, the heavy notation is not possible. Thus, we have worked on improving more precise wording and removing sections or wording that is distracting or not needed. Thank the Editor and Associate Editor for helping us improve upon what we believe to be an important paper and breakthrough in the literature.

%% I would not make a large list of all the notation we have changed as it may cause the Editor to go picking over the paper more.

\reply{
\textcolor{teal}{We thank the Editor for the opportunity to revise our paper and for the suggestions. We have improved notation, when possible, however, as noted by the Editor in certain places, such as in the computational speeds ups of Section 4, this is difficult to avoid. We have done our utmost to simplify notation (when possible) and improve the parsing of sentences. Finally, we have attempted to clarify our overall contributions to the literature, which we highlight below. In addition, we provide responses to all the Editor's and Associate Editor's comments below.}

We propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}. This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}. We employ the decision theoretic  technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite. This discovery allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs via a novel method, storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file. Finally, we provide two real illustrations of our methodology and simulation studies with reproducible software.

@@

	``Consider a data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$". See Section 2, page 3. 
}

\setcounter{responsectr}{0}
\section{Reviewer Comments}

	\pointRaised{R}{%
	For the missing data treatment now included just before Section 2.1, is the assumption truly missing at random, or missing completely at random? Or, does this distinction not matter because of the independence assumed across elements of the comparison vector?}

	\reply{\textcolor{teal}{The distinction does not matter due to assuming that the comparison vectors are conditionally independent given the coreference matrix (or matching label). We now state the following in Section 3, page 5:}

	\textcolor{teal}{``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR). Due the the conditional independence assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}

%	Furthermore, both MAR and MCAR lead to assuming the missing data is ignorable in this situation, which implies that the marginalized likelihood under all three assumptions is the same."

%	``Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}). With either the ignorability or MAR assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}. Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p. 12]{LittleRubin2002}). To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation."
%

%	(MAR, per \cite{LittleRubin2002}). With either the ignorability or MAR assumption, we can marginalize over the missing data in (6a) and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}. Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p. 12]{LittleRubin2002}). To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation."
	}

@@

	%% RCS: See two comments below regarding MAR versus MCAR. I have an alternative response for the one regarding miss at random.

%	\textcolor{teal}{Answer for MAR: Yes, the assumption is truly missing at random. This distinction between truly missing at random and missing completely at random does not matter due to two assumptions. Following Sadinle (2017), we assume that the fields in the comparison vectors are conditionally independent given the coreference matrix and we assume that the missing comparisons are ignorable (Sadinle (2017). Therefore, our likelihood is the same as in Sadinle (2017) (see equations (3) and (4) of his paper). As stated by Sadinle (2017), ``we work under the assumption of ignorability of the missingness mechanism for the comparisons so that we can base our inferences on the marginal distribution of the observed comparisons (Little and Rubin, p. 90)."  We refer to Sadinle (2017), Section 4.2, page 604 for further details.}

%	\reply{The assumption is missing completely at random (MCAR) due to fact that we assume the fields in the comparison vectors are conditionally independent given the coreference matrix (or matching label). The assumption is not missing at random (MAR) as we do not assume that the missing comparisons are also ignorable. See Section 3, page 5 regarding clarification of this point \citep{LittleRubin2002}.}
%	\textcolor{red}{Still checking on this and looking at Little and Rubin.}

### 2023-07-20 ba_resubmit/kundinger_resubmit2.tex
address fix that patrick ball asked for over email.

\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors. To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}). With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}. Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p. 12]{LittleRubin2002}). To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation.}

@@

	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017}. Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors. In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study, a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.

@@

	The country of El Salvador was immersed in civil war from 1980 to 1991. We are interested in estimating the total number of casualties from the war. We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL) and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish).\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization \citep{howland2008rescate, ball2000salvadoran}. The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable. When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists. Instead, record linkage techniques must be used to merge data files before analyzing the data \citep{lum2013applications}.

	There are several challenges with these data. First, both data files have been automatically digitized, which inherently leads to some degree of typographical error. Second, the only fields recorded are given name, last name, date of death, and place of death. It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.

>>>>

\textcolor{teal}{We now explain how we accommodate missing fields, which leads to missing comparison vectors. To denote a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are  missing at random (MAR) or missing completely at random (MCAR). Due the the conditional independence assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}) and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002}, \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}.}

%Assume the comparison vectors are conditionally independent given the coreference matrix (or matching label) and that missing comparisons are ignorable or missing at random (MAR, per \cite{LittleRubin2002}). With either the ignorability or MAR assumption, we can marginalize over the missing data in (\ref{eqn:likelihood}), and do all computation simply using the observed data. For details, refer to \cite[Section 6.2]{LittleRubin2002} and \cite[Section 3.1]{sadinle_detecting_2014}, and \citep[Section 4.2]{sadinle_bayesian_2017}. Due to the conditional independence assumption, the missingness mechanism is equivalent to being missing completely at random (MCAR, per \cite[p. 12]{LittleRubin2002}). To summarize, we can assume the missingness is ignorable, MAR, or MCAR in this situation.}

@@

	In our first case study, we revisit data from the El Salvadoran Civil War analyzed by \cite{sadinle_bayesian_2017} and others \citep{lum2013applications, jewell2018accounting}. Though the data files used in this case study are small, it shows how the computational complexity of \texttt{fabl} depends on the number of unique agreement patterns found in the data, and how significant computational gains can be achieved by simplifying the construction of the comparison vectors. In the second case study, we apply \texttt{fabl} to link records from the National Long Term Care Study, a larger linkage task that is not feasible in reasonable time under \texttt{BRL} with typical computing setups.

@@

	The country of El Salvador was immersed in civil war from 1980 to 1991. We are interested in estimating the total number of casualties from the war. We utilize lists of casualties from the war, one collected by El Rescate - Tutela Regal (ERTL)\citep{howland2008rescate} and another from the Salvadoran Human Rights Commission (CDHES, by its acronym in Spanish)\citep{ball2000salvadoran}.\footnote{We thank the Human Rights Data Analysis Group (HRDAG) for granting access to these data.} The ERTL dataset comprises digitized denunciations published throughout the conflict, and the CDHES dataset comprises casualties reported directly to the organization. The ERTL required additional investigation before recording denunciations as human rights abuses, and reports to the CHDES were made shortly after the events occurred; thus, both data files are thought to be fairly reliable. When estimating the total number of casualties, one cannot simply sum the numbers recorded by each organization, as it is likely that the same individuals are recorded in multiple casualty lists. Instead, record linkage techniques must be used to merge data files before analyzing the data \citep{lum2013applications}.

	There are several challenges with these data. First, both data files have been automatically digitized, which inherently leads to some degree of typographical error. Second, the only fields recorded are given name, last name, date of death, and place of death. \textcolor{teal}{While it may be common in other data sets for a parent and child to share the same given name, in this data set, it is more common for cousins to share this information, especially in villages, where there may be only four to five extended families. In Latin America, an individual receives a surname from their father and their mother, leading to a limited number of surnames in small villages.}
	%% I believe that Patrick said that this is not common in this data set
%	It is relatively common for a parent and child to share the same given name, resulting in indistinguishable records for two different individuals.
